---
title: "Modelstroke"
author: "Utkarshna Sinha"
date: '2021-04-21'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE}
library(readr)
library(tidyverse)
library(caret)
library(ROSE)
library(modelr)
library(glmnet)
library(kknn)
setwd("~/Desktop/DS5110/Project/")
all_data <- read.csv("./train_strokes.csv", header = TRUE)
mean_bmi = mean(all_data$bmi,na.rm = T)

all_data = all_data %>%
              mutate(bmi = ifelse(is.na(bmi),mean_bmi,bmi))
stroke_data2 <- all_data %>% na.omit() %>% 
  mutate(age_group=case_when(age >= 91 & age <= 109 ~ '90+',
                             age >= 80 & age <= 90 ~ '80 - 89',
                            age >= 70 & age <= 79 ~ '70 - 79',
                            age >= 60 & age <= 69 ~ '60 - 69',
                            age >= 50 & age <= 59 ~ '50 - 59',
                            age >= 40  & age <= 49 ~ '40 - 49', 
                            age >= 30  & age <= 39 ~ '30 - 39', 
                            age >= 19  & age <= 29 ~ '19 - 29',
                            age >= 0  & age <= 18 ~ '0 - 18',))
```

```{r}
stroke_data2$age_group <- as.factor(as.character(stroke_data2$age_group))
stroke_data2$stroke <- as.factor(as.integer(stroke_data2$stroke))
stroke_data2$age <- as.factor(as.numeric(stroke_data2$age))
stroke_data2$gender <- as.factor(as.character(stroke_data2$gender))
stroke_data2$ever_married <- as.factor(as.character(stroke_data2$ever_married))
stroke_data2$work_type <- as.factor(as.character(stroke_data2$work_type))
stroke_data2$Residence_type <- as.factor(as.character(stroke_data2$Residence_type))
stroke_data2$smoking_status <- as.factor(as.character(stroke_data2$smoking_status))
stroke_data2$hyptertension <- as.factor(as.integer(stroke_data2$hypertension))
stroke_data2$heart_disease <- as.factor(as.integer(stroke_data2$heart_disease))
```

```{r warning=FALSE}
# oversample and undersample with caret 
# sensitivity & specificity be more balanced at least 0.5 - 1
# manually sample to generate indices (replace=TRUE) with sample, undersample with replace=FALSE

index <- createDataPartition(stroke_data2$stroke, p = 0.8, list = FALSE)
stroke_t <- stroke_data2[as.integer(index),]
stroke_test <- stroke_data2[-as.integer(index),]

#table(stroke_data2$stroke[index])
```

ROSE generates synthetic balanced samples and thus allows to strenghten the subsequent estimation of any binary classifier.

```{r}
stroke_train <- ROSE(stroke ~ ., data = stroke_t, seed=22)$data
t3<-table(stroke_train$stroke)
```

Create ROSE graph
```{r}
t3 <- as.data.frame(t3) %>% 
  mutate(Stroke=recode(Var1, "1" ="Had Stroke", "0" ="No Stroke" )) 
ggplot(t3, aes(x=Stroke, y=Freq, fill=Stroke)) +
  geom_bar(stat="identity") +
  labs(x = "Stroke", y = "Frequency",
       title="Class distribution for Training Set after ROSE sampling", fill="Stroke") +
  theme_minimal()
```

```{r}
set.seed(99)
stroke_fit1 <- glm(stroke ~ age, 
                   data=stroke_train, 
                   family=binomial(link="logit"))

s_prob2 <- predict(stroke_fit2, newdata=stroke_test, type="response")

#hist(s_prob2)
s_pred2 <- ifelse(s_prob2 > 0.5, 1, 0)
mean(s_pred2 == stroke_test$stroke, na.rm = TRUE)
table(s_pred2, stroke_test$stroke)
```

```{r warning=FALSE}
set.seed(78)
s_fit1 <- train(stroke ~ age, data=stroke_train,
              method="glm", family=binomial(link="logit"),
              preProcess="medianImpute",
              trControl=trainControl(method="none"),
              na.action=na.pass)

confusionMatrix(predict(s_fit1, stroke_test, na.action=na.pass),
                 stroke_test$stroke)
```

```{r}
summary(s_fit1)
```

```{r warning=FALSE}
set.seed(78)
s_fit2 <- train(stroke ~ age + hyptertension , data=stroke_train,
              method="glm", family=binomial(link="logit"),
              preProcess="medianImpute",
              trControl=trainControl(method="none"),
              na.action=na.pass)

confusionMatrix(predict(s_fit2, stroke_test, na.action=na.pass),
                 stroke_test$stroke)
```

```{r warning=FALSE}
set.seed(78)
s_fit3 <- train(stroke ~ age + hyptertension + heart_disease , data=stroke_train,
              method="glm", family=binomial(link="logit"),
              preProcess="medianImpute",
              trControl=trainControl(method="none"),
              na.action=na.pass)

confusionMatrix(predict(s_fit3, stroke_test, na.action=na.pass),
                 stroke_test$stroke)
```

```{r}
summary(s_fit3)
```


```{r warning=FALSE}
set.seed(48)
s_fit4 <- train(stroke ~ age + hyptertension + heart_disease + avg_glucose_level, data=stroke_train,
              method="glm", family=binomial(link="logit"),
              preProcess="medianImpute",
              trControl=trainControl(method="none"),
              na.action=na.pass)

confusionMatrix(predict(s_fit4, stroke_test, na.action=na.pass),
                 stroke_test$stroke)
```

 Does not run
```{r warning=FALSE}
set.seed(78)
s_fit5 <- train(stroke ~ ., data=stroke_train,
              method="glm", family=binomial(link="logit"),
              preProcess="medianImpute",
              trControl=trainControl(method="none"),
              na.action=na.pass)

s_fit5
```

## Plotting the ROC

Changing the cutoff probability for class assignment can affect the sensitivity and specificity. It can be useful to calculate and plot the tradeoff between sensitivity and specificity for different cutoffs. This is traditionally visualized as an ROC curve, which plots the *true positive rate* (sensitivity) against the *false positive rate* (1 - specificity): First we write functions for calculating sensitivity and specificity for a given probability cutoff.

```{r}
sens <- function(c, p, ref, positive = levels(ref)[2])
{
  mean((p > c)[ref == positive], na.rm=TRUE)
}
sens(0.5, s_prob, stroke_test$stroke)
sens(0.5, s_prob2, stroke_test$stroke)
sens(0.5, s_prob3, stroke_test$stroke)

spec <- function(c, p, ref, negative = levels(ref)[1])
{
  mean((p < c)[ref == negative], na.rm=TRUE)
}
spec(0.5, s_prob, stroke_test$stroke)
spec(0.5, s_prob2, stroke_test$stroke)
spec(0.5, s_prob3, stroke_test$stroke)
```

```{r}
roc <- tibble(p=seq(from=0, to=1, by=0.01)) %>%
  mutate(sensitivity = map_dbl(p, sens, p=s_prob3, ref=stroke_test$stroke),
         specificity = map_dbl(p, spec, p=s_prob3, ref=stroke_test$stroke),
         TPR=sensitivity,
         FPR=1 - specificity)
roc

ggplot(roc, aes(x=FPR, y=TPR)) + 
  geom_path(color="red", size=1) +
  geom_vline(xintercept=0, color="green", linetype="dotdash") +
  geom_hline(yintercept=1, color="green", linetype="dotdash") +
  geom_abline(intercept=0, slope=1, color="blue", linetype="dotted") +
  labs(x="False positive rate (1 - specificity)",
       y="True positive rate (sensitivity)") +
  theme_minimal()
```





```{r}
set.seed(888)
knn_fit <- train.kknn(stroke ~ age + hyptertension + heart_disease + avg_glucose_level, 
                 data=stroke_train, na.action = na.pass)
              
summary(knn_fit)
```
```{r}
knn_pred <- predict(knn_fit, newdata=stroke_test)

mean(knn_pred == stroke_test$stroke, na.rm = TRUE)
table(knn_pred, stroke_test$stroke)
knn_fit
```


```{r}
confusionMatrix(predict(knn_fit, stroke_test, na.action=na.pass), stroke_test$stroke)
```


```{r}
set.seed(88)
knn_fit2 <- train.kknn(stroke ~ age + hyptertension + heart_disease, 
                 data=stroke_train, na.action = na.pass)

confusionMatrix(predict(knn_fit2, stroke_test, na.action=na.pass), stroke_test$stroke)
```

```{r}
set.seed(48)
knn_fit3 <- train.kknn(stroke ~ age + hyptertension, 
                 data=stroke_train, na.action = na.pass)

confusionMatrix(predict(knn_fit3, stroke_test, na.action=na.pass), stroke_test$stroke)
```

```{r}
set.seed(48)
knn_fit4 <- train(stroke ~ age , method="knn",
                 data=stroke_train, na.action = na.pass)

confusionMatrix(predict(knn_fit4, stroke_test, na.action=na.pass), stroke_test$stroke)
```

```{r}
set.seed(48)
knn_fit5 <- train.kknn(stroke ~ age + avg_glucose_level, 
                 data=stroke_train, na.action = na.pass)

confusionMatrix(predict(knn_fit5, stroke_test, na.action=na.pass), stroke_test$stroke)
```

```{r}
confusionMatrix(predict(knn_fit2, stroke_test, na.action=na.pass), stroke_test$stroke)
```


GLM without caret
Do the glm fit for logit
```{r}
set.seed(8)
stroke_fit <- glm(stroke ~ age, data=stroke_train, family=binomial(link="logit"))
s_prob <- predict(stroke_fit, newdata=stroke_test, type="response")
hist(s_prob)
```

Now we can calculate the accuracy:

```{r}
set.seed(8)
s_pred <- ifelse(s_prob > 0.5, 1, 0)

mean(s_pred == stroke_test$stroke, na.rm = TRUE)
```

And a confusion matrix:

```{r}
table(s_pred, stroke_test$stroke)
confusionMatrix()
```

```{r}
set.seed(99)
stroke_fit2 <- glm(stroke ~ age+hypertension, 
                   data=stroke_train, family=binomial(link="logit"))
s_prob2 <- predict(stroke_fit2, newdata=stroke_test, type="response")

#hist(s_prob2)
s_pred2 <- ifelse(s_prob2 > 0.5, 1, 0)
mean(s_pred2 == stroke_test$stroke, na.rm = TRUE)
table(s_pred2, stroke_test$stroke)
```

```{r}
set.seed(88)
stroke_fit3 <- glm(stroke ~ age+hypertension+heart_disease, 
                   data=stroke_train, family=binomial(link="logit"))
s_prob3 <- predict(stroke_fit3, newdata=stroke_test, type="response")

#hist(s_prob3)
s_pred3 <- ifelse(s_prob3 > 0.5, 1, 0)
mean(s_pred3 == stroke_test$stroke, na.rm = TRUE)
table(s_pred3, stroke_test$stroke)
confusionMatrix(predict(stroke_fit3, stroke_test, na.action=na.pass),
                stroke_test$stroke)
```
```{r}
set.seed(83)
stroke_fit4 <- glm(stroke ~ age+hypertension+heart_disease +avg_glucose_level, 
                   data=stroke_train, family=binomial(link="logit"))
s_prob4 <- predict(stroke_fit4, newdata=stroke_test, type="response")

#hist(s_prob4)
s_pred4 <- ifelse(s_prob4 > 0.5, 1, 0)
mean(s_pred4 == stroke_test$stroke, na.rm = TRUE)
table(s_pred4, stroke_test$stroke)
```
```{r}
set.seed(23)
stroke_fit5 <- glm(stroke ~ age+hypertension+heart_disease +bmi, 
                   data=stroke_train, family=binomial(link="logit"))
s_prob5 <- predict(stroke_fit5, newdata=stroke_test, type="response")

#hist(s_prob4)
s_pred5 <- ifelse(s_prob5 > 0.5, 1, 0)
mean(s_pred5 == stroke_test$stroke, na.rm = TRUE)
table(s_pred5, stroke_test$stroke)
```